Lesson 6
========================================================

### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
# Let's start by examining two variables in the data set.
# The scatterplot is a powerful tool to help you understand the relationship between two continuous variables.

# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond characteristics to help us figure out whether the price advertised for any given diamond is reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight. Create a scatterplot of price (y) vs carat weight (x). Limit the x-axis and y-axis to omit the top 1% of values

library(ggplot2)
diamonds <- diamonds
ggplot(aes(x = carat, y = price), data = diamonds) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, .99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, .99))) +
  geom_point(alpha = 0.3, fill = I('#F79420'), color = I('black'), shape = 21) +
  stat_smooth(method='lm')
```

***

### Price and Carat Relationship
Response:
In general, the price increases as the carat of diamonds increase, but there is a big range of prices for the same carat, which might be because that other factors might affect the diamond price as well.
Non-linear relationship, maybe exponential
***

### Frances Gerety
Notes:

#### A diamonds is


***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:

```{r ggpairs Function}
# # install these if necessary
# # To plot each variable against each other
# install.packages('GGally')
# # For a variety of things
# install.packages('scales')
# # To summarize the regression
# install.packages('memisc')
# # For few other things
# install.packages('lattice')
# # For various functions
# install.packages('MASS')
# # To recode variables
# install.packages('car')
# # To reshape and wrangle the data
# install.packages('reshape')
# # To create interesting summaries and transmissions
# install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
# ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
ggsave("diamonds.png")
#,plot = last_plot()
```

What are some things you notice in the ggpairs output?
Response:
No matter what the 4c levels are, most diamond prices are in the low range.
There might be relationships between price vs. clarity, price vs. color
***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# We've put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

# You can self-assess your work with the plots
# in the solution video.

# ALTER THE CODE BELOW THIS LINE
# ==============================================

library(gridExtra)

plot1 <- qplot(x = price, data = diamonds, binwidth = 100, fill = I('#099DD9')) + 
  ggtitle('Price')

plot2 <- qplot(x = price, data = diamonds, binwidth = 0.01, fill = I('#F79420')) +
  ggtitle('Price (log10)') + 
  scale_x_log10()

grid.arrange(plot1, plot2, ncol = 2)
```

***

### Connecting Demand and Price Distributions
Notes:
There are two peaks in the histogram of diamond price, at around 1000 and 9000, which means most people have their budgets around these two points.
***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(aes(x = carat, y = price), data = diamonds) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, .99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, .99))) +
  scale_y_log10() +
  geom_point(alpha = 0.3, fill = I('#F79420'), color = I('black'), shape = 21) +
  stat_smooth(method='lm')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```


```{r Overplotting Revisited}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.

# If you need hints, see the Instructor Notes.
# There are three hints so scroll down slowly if
# you don't want all the hints at once.

# ALTER THE CODE BELOW THIS LINE
# =======================================================================
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 0.75, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

```{r Price vs. Carat and Clarity}
# A layer called scale_color_brewer() has 
# been added to adjust the legend and
# provide custom colors.

# See if you can figure out what it does.
# Links to resources are in the Instructor Notes.

# You will need to install the package RColorBrewer
# in R to get the same colors and color palettes.

# =======================================
library(RColorBrewer)
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:
Yes. For diamonsd with same weight, the price increases as the clarity of the diamonds gets better.
***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
# A layer called scale_color_brewer() has 
# been added to adjust the legend and
# provide custom colors.

# See if you can figure out what it does.
# Links to resources are in the Instructor Notes.

# You will need to install the package RColorBrewer
# in R to get the same colors and color palettes.

# =======================================
library(RColorBrewer)
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:
Cut does not hava a big impact on the diamond price, since there is no clear difference on price for diamonds with the same weight but different cut levels.
***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
# Finally, let's use diamond color to color our plot.

# Adjust the code below to color the points by diamond colors
# and change the titles.

ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = F,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:
Yes. For diamonsd with same weight, the price increases as the color of the diamonds gets better.
***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:
Use I wrapper around each of the variables.
I stands for as is.
In this case, it tells R to use the expression inside the I function to transform a variable before using it in the regression. This is instead of instructing R to interpret these symbols as part of the formula to construct the design matrix for the regression.
```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
library(memisc)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should spend 10-20 min on this)

Response:
This model does not include the interaction within the predictors in the regression model.
Other than carat, color also have a big impact on the price outcome, but it is not included as an significant variable in the model.
This model has included from 1st power to the 7th power for cut, color, and clarity, which is not necessary.

2008-2014
- We should count for inflation
- 2008 global recession (global diamond report)
- Diamond Market in China heating up
- Uneven recovery/price increase across different carat weight

***

### A Bigger, Better Data Set
Notes:
If the code shown in the video does not work, the dataset is available for download here: https://github.com/solomonm/diamonds-data. Click on the BigDiamonds.Rda link, then click on the "Raw" button to start the download. Once downloaded, the data can be loaded using the command load("BigDiamonds.rda"), assuming that the file is in your working directory.

You can learn how to scrape data from the web just like Solmon did! Take Data Wrangling with MongoDB: Data Manipulation and Retrieval with us.
```{r A Bigger, Better Data Set}
#install.package('bitops')
#install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
# load(rawConnection(diamondsurl))
load("BigDiamonds.rda")
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
# Your task is to build five linear models like Solomon did for the diamonds data set only this time you'll use a sample of diamonds from the diamondsbig data set.

# Be sure to make use of the same variables (logprice, carat, etc.) and model names (m1, m2, m3, m4, m5).

# To get the diamondsbig data into RStudio on your machine, copy, paste, and run the code in the Instructor Notes. There's 598,024 diamonds in this data set!

# Since the data set is so large, you are going to use a sample of the data set to compute the models. You can use the entire data set on your machine which will produce slightly different coefficients and statistics for the models.

# This exercise WILL BE automatically graded.

# You can leave off the code to load in the data. We've sampled the data for you. You also don't need code to create the table output of the models. We'll do that for you and check your model summaries (R^2 values, AIC, etc.)

# Your task is to write the code to create the models.

# DO NOT ALTER THE CODE BELOW THIS LINE (Reads in a sample of the diamondsbig data set)
#===========================================================================================
diamondsBigSample <- read.csv('diamondsbig.csv')


# ENTER YOUR CODE BELOW THIS LINE. (Create the five models)
#===========================================================================================

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsBigSample[diamondsBigSample$price<10000 & diamondsBigSample$cert == 'GIA',])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)



# DO NOT ALTER THE CODE BELOW THIS LINE (Tables your models and pulls out the statistics)
#===========================================================================================
suppressMessages(library(lattice))
suppressMessages(library(MASS))
suppressMessages(library(memisc))
models <- mtable(m1, m2, m3, m4, m5)
```

## Confidence Interval
The prediction interval here may be slightly conservative, as the model errors are heteroskedastic over carat (and hence price) even after our log and cube-root transformations.

See the output of the following code.

How could we do better? If we care most about diamonds with carat weights between 0.50 and 1.50, we might restrict the data we use to fit our model to diamonds that are that size - we have enough data
```{r}
dat = data.frame(m4$model, m4$residuals)

with(dat, sd(m4.residuals))

with(subset(dat, carat > .5 & carat < 1.5), sd(m4.residuals))

dat$resid <- as.numeric(dat$m4.residuals)
ggplot(aes(y = resid, x = round(carat, 2)), data = dat) +
  geom_line(stat = "summary", fun.y = sd)
```

***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you've loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
modelEstimate
exp(modelEstimate)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

## Predictions

Example Diamond from BlueNile:
Round 1.25 Ideal E VVS2 $11,713

```{r}
#Be sure you've loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.25, cut = "Ideal",
                         color = "E", clarity="VVS2")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
modelEstimate
exp(modelEstimate)
```

```{r}
#Be sure you've loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 0.62, cut = "Ideal",
                         color = "I", clarity="VS2")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
modelEstimate
exp(modelEstimate)
```
Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

